---
title: "Course Projet"
output: html_notebook
---

# Load Libraries (tibbles tibbles tibbles)

```{r}
# libraries needed readr, dplyr, tidyr, tibbles
library(readr) # reading tables
library(dplyr) # manipulation
library(tidyr)
library(tibble)
```

# Create tmp dir for file downloads

```{r}
project_temp <- tempdir()
if(!dir.exists(project_temp)){
    dir.create(project_temp)
}

data_url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
data_arch <- paste0(project_temp, "\\data.zip")
download.file(data_url, destfile = data_arch)
unzip(data_arch, exdir = project_temp, junkpaths = T)
```

# Read files needed for tidying

```{r}
features <- readr::read_table(
    paste0(project_temp,"//features.txt"),
    col_names = c("num", "feat"),
    col_types = "ic")
activities <- readr::read_table(
    paste0(project_temp,"//activity_labels.txt"),
    col_names = c("num", "activity"),
    col_types = "ic")
subject_test <- read_table(
    paste0(project_temp,"//subject_test.txt"),
    col_names = c("subject"),
    col_types = "i")
subject_train <- read_table(
    paste0(project_temp,"//subject_train.txt"),
    col_names = c("subject"),
    col_types = "i")
x_test <- read_table(
    paste0(project_temp,"//X_test.txt"),
    col_names = features$feat)
x_train <- read_table(
    paste0(project_temp,"//X_train.txt"),
    col_names = features$feat)
y_test <- read_table(
    paste0(project_temp,"//y_test.txt"),
    col_names = "code")
y_train <- read_table(
    paste0(project_temp,"//y_train.txt"),
    col_names = "code")
```

# Merging Files

1.  combine the x files

2.  combine the y files

3.  combine subject files

4.  merge alld the columns from X, Y and subjects

    ```{r}
    X <- rbind(x_train, x_test)
    Y <- rbind(y_train, y_test)
    Subject <- rbind(subject_train, subject_test)
    Merged_Data <- cbind(Subject, Y, X)
    ```

# Extract columns

Extracting columns (variables) subject, code, any other column containing "mean" or "std" and saving it in my Tidies up data table td.

```{r}
td <- Merged_Data %>%
    select(subject, code, contains("mean"), contains("std"))
```

# Using Names instead of codes as the activity

-   too many ways to do the same thing, making my standard to always use tibbles. Any other recommendations are greatly appreciated.

```{r}
td <- as_tibble(td)
td$code <- activities[td$code, 2]
```

# Renaming columns

substituting abbreviations with full words

```{r}
names(td)[2] = "activity"
names(td)<-gsub("Acc", "Accelerometer", names(td))
names(td)<-gsub("Gyro", "Gyroscope", names(td))
names(td)<-gsub("BodyBody", "Body", names(td))
names(td)<-gsub("Mag", "Magnitude", names(td))
names(td)<-gsub("^t", "Time", names(td))
names(td)<-gsub("^f", "Frequency", names(td))
names(td)<-gsub("tBody", "TimeBody", names(td))
names(td)<-gsub("-mean()", "Mean", names(td), ignore.case = TRUE)
names(td)<-gsub("-std()", "STD", names(td), ignore.case = TRUE)
names(td)<-gsub("-freq()", "Frequency", names(td), ignore.case = TRUE)
names(td)<-gsub("angle", "Angle", names(td))
names(td)<-gsub("gravity", "Gravity", names(td))
```

# Creating Summary Data

using summarize on the groups specified then writing it to the project output folder.

```{r}
summary_data <- td %>%
    group_by(subject, activity) %>%
    summarise_all(list(mean))
write.table(FinalData, "summary_data.txt", row.names = F )
```

# Removing Temporary File/Folder

```{r}
unlink(project_temp, recursive = T, force = T)
file.remove(project_temp, recursive = T)
rm(list = ls())
detach(readr)
detach(dplyr)
detach(tidyr)
detach(tibble)
```
